---
title: "Chapter Two Workflow"
author: "Fiona Spooner"
date: "19 October 2017"
output: html_document
---

####Europe####

Have ignored the bootstrapping of the models and the use of max, min and precipitation data, and the dredging!
```{r, warning=FALSE, message=FALSE}
LPI<-read.csv("LPI_pops_20160523_edited.csv")

LPI_pop<-subset(LPI, Specific_location==1 & System !="Marine" & Class != "Actinopterygii"& Class != "Cephalaspidomorphi"& ID != 4438)

ID<-LPI_pop$ID
pop_data<- LPI_pop[,c(1,65:127)]

pop_datab <- (pop_data [,2:64] !="NULL")
points_per_pop1950_2012 = rowSums(pop_datab)
length_id <- data.frame(ID,points_per_pop1950_2012)

LPI_EU<-merge(length_id, LPI_pop, by = "ID")
LPI_EU<-subset(LPI_EU, points_per_pop1950_2012 >=2)

LPI_EU2<-LPI_EU[,c(1:3,66:128)]   #just pop trends
LPI_EU2[LPI_EU2 == 'NULL'] = NA

```

```{r, message = FALSE, warning = FALSE}
library(taRifx)
library(mgcv)
library(zoo)

doFit = function(sp_name) {
  spid2 = subset(LPI_EU2, ID == sp_name)   #subsetting the population data by each population 
  spid = spid2[,4:66]                     #subsetting only the dates
  colnames(spid)<-1950:2012              #renaming the date column names as R doesn't like numbered column names
  
  name<-spid2$Binomial
  id<-spid2$ID
  points<-spid2$points_per_pop1950_2012
  name_id<-paste(name, id, sep="_") #creating id for naming files of plots
  Date<-as.numeric(colnames(spid))
  spidt<-destring(t(spid))
  time<-length(min(which(!is.na(spidt))):max(which(!is.na(spidt))))
  missing<-time-points
  
  Year<-Date[min(which(!is.na(spidt))):max(which(!is.na(spidt)))]
  Population<-spidt[min(which(!is.na(spidt))):max(which(!is.na(spidt)))]
  Population[Population == 0] <- mean(Population, na.rm=TRUE)*0.01 
  
  df<-data.frame(Year,Population)
    
  
  if (points >=6) {           ###should I be trying GAMs for populations with less than six points and if that doesn't fit then use a linear model - here I am automatically fitting a linear model if there are less than six points
    PopN = log10(Population)
    if (length(na.omit(PopN)) >=6) {
      SmoothParm = round(length(na.omit(PopN))/2)    ####added na.omit in as was getting " A term has fewer unique covariate combinations than specified maximum degrees of freedom" error
    } else{
      SmoothParm=3
    }
    mg2<-mgcv:::gam(PopN ~ s(Year, k=SmoothParm), fx=TRUE)
    pv2 <- predict(mg2,df,type="response",se=TRUE) 
    R_sq2<-summary(mg2)$r.sq
    model<-1
    pv2$fit[pv2$fit <= 0] <- NA
    lambda2<-diff(pv2$fit)
    lambda_sum2<-sum(lambda2, na.rm=TRUE)
    lambda_mean2<-mean(lambda2, na.rm=TRUE)
 
  } else {
    SmoothParm<-NA
    PopN = log10(Population)
    ml2<-lm(PopN~df$Year)
    R_sq2<-summary(ml2)$r.sq
    model<-0
    Pop_interp2<-na.approx(PopN)
    Pop_interp2[Pop_interp2<=0] <- NA
    lambda2<-diff(Pop_interp2)
    lambda_sum2<-sum(lambda2, na.rm=TRUE)
    lambda_mean2<-mean(lambda2, na.rm=TRUE)
  }
  
  res_df = data.frame(sp_name=sp_name, points=points, SmoothParm=SmoothParm, r_sq=R_sq2, model=model,lambda_sum=lambda_sum2,lambda_mean=lambda_mean2,time=time, missing=missing)
  #print(id)
  return(res_df)
}


all_df_list <- lapply(unique(LPI_EU2$ID), doFit)


```

```{r}
all_matrix <- matrix(unlist(all_df_list, use.names =FALSE), ncol=9, byrow=TRUE)
all_df <- data.frame(all_matrix)
colnames(all_df) <- c("ID", "points","SmoothParm", "r_sq", "model", "lambda_sum","lambda_mean", "length_time", "missing_years")
wd<-getwd()
```

```{r}
write.csv(all_df, paste(wd,"/Chapter_Two/Europe/", "Global_Population_Lambdas.csv", sep=""))

```

Download climate data from here - http://www.ecad.eu/download/ensembles/download.php

```{r, eval=FALSE}

rmean<-brick("C:/Users/Fiona/Desktop/PhD/Climate/Europe/MeanT/tg_0.25deg_reg_v11.0.nc", varname = "tg")

tm <- seq(as.Date('1950-01-01'), as.Date('2014-12-31'), 'day')
r2 <- setZ(rmean, tm, 'days')
xmean <- zApply(r2, by=as.yearmon, fun=mean, name='months')   #this bit can take ages (hours) - it is summarasing the daily data into monthly data

writeRaster(xmean, filename="Europe_mean_mon.grd", bandorder='BIL', overwrite=TRUE)

```


```{r}

library(raster)

xy<-cbind(LPI_EU$Longitude, LPI_EU$Latitude)

id<-LPI_EU$ID

binomial<-LPI_EU$Binomial

xmean<- brick(paste(wd,"/Chapter_Two/Europe/", "Europe_mean_mon.grd", sep=""))

```

```{r,eval=FALSE}
all_EU_mean_data<-data.frame(id=numeric(0), Binomial=character(0),xy=numeric(0), year=numeric(0), month=numeric(0), rasterex=numeric(0))


for (i in 1:nlayers(xmean)) {

  rasterex<- extract(xmean[[i]], xy, buffer=50000, fun=mean, na.rm=TRUE)
  date<-names(xmean[[i]])
  date2 <- as.yearmon(date, "%b.%Y")
  dataex<-data.frame(id, binomial,xy, format(date2, "%Y"),format(date2, "%b"), rasterex)   
 # print(date2)
  all_EU_mean_data = rbind(all_EU_mean_data, dataex)
  
}

```

```{r, eval=FALSE}
colnames(all_EU_mean_data)[c(2:7)]<-c("Binomial","Longitude","Latitude", "Year", "Month", "Mean_Temp")

all_EU_mean_data_na_omit<-na.omit(all_EU_mean_data)

write.csv(all_EU_mean_data, paste(wd,"/Chapter_Two/Europe/" ,"Mean_Temp_E-OBS_50k_buff.csv", sep=""))

write.csv(all_EU_mean_data_na_omit, paste(wd,"/Chapter_Two/Europe/" ,"Mean_Temp_E-OBS_50k_buff_na_omitted.csv", sep=""))


```


```{r}

all_EU_mean_data_na_omit<-read.csv(paste(wd,"/Chapter_Two/Europe/" ,"Mean_Temp_E-OBS_50k_buff_na_omitted.csv", sep=""))


```

```{r, warning=FALSE}
library(plyr)
library(broom)

LPI_EU2<-LPI_EU[LPI_EU$ID %in% all_EU_mean_data_na_omit$id & LPI_EU$ID!=18236 &  LPI_EU$ID!=18239 ,]
nrow(LPI_EU2)

doMean = function(sp_name) {
  spid2 = subset(LPI_EU2, ID == sp_name)   #subsetting the population data by each population
  spid = spid2[,66:128]                     #subsetting only the dates
  colnames(spid)<-1950:2012              #renaming the date column names as R doesn't like numbered column names
  climid=subset(all_EU_mean_data_na_omit, id == sp_name)  #subsetting the climate data by each population
  
  year_temp <- ddply(climid, "Year", summarise,          #calculating the annual mean for max temp, min temp and precipitation
                     mean_mean = mean(na.omit(Mean_Temp)))
  
  lt_avg <- ddply(climid, "id", summarise,              #calculating the long term mean (1950-2014) of the max temp, min temp and precipitation - for each population
                  lt_avg_mean = mean(na.omit(Mean_Temp)))
  
  year_temp$anom_mean = year_temp$mean_mean - lt_avg$lt_avg_mean     #calculating anomalies 1950-2014
  
  
  name<-spid2$Binomial
  id<-spid2$ID
  points<-spid2$points_per_pop1950_2012
  name_id<-paste(name, id, sep="_") #creating id for naming files of plots
  Date<-as.numeric(colnames(spid))
  spidt<-destring(t(spid))
  
  Year<-Date[min(which(!is.na(spidt))):max(which(!is.na(spidt)))]
  Population<-spidt[min(which(!is.na(spidt))):max(which(!is.na(spidt)))]
  
  Mean_mon<-climid[climid$Year %in% Year, ]$Mean_Temp

  Mean_anom<-year_temp$anom_mean[min(which(!is.na(spidt))):max(which(!is.na(spidt)))]
  Mean<-year_temp$mean_mean[min(which(!is.na(spidt))):max(which(!is.na(spidt)))]

  if (sum(is.nan(Mean))!=length(Mean)){
  
  lm_mean<-lm(Mean~Year)
  lm_mean_df<-tidy(lm_mean)[2,]  
  mean_df<-cbind(id,lm_mean_df)
  
  } else{
    
  mean_df<-matrix(c(id,NA,NA,NA,NA,NA), nrow=1, ncol=6)
  colnames(mean_df)<-c("id", "term", "estimate", "std.error", "statistic", "p.value")
  mean_df<-data.frame(mean_df)
  }
  

  #print(id)  
  return(mean_df)
}

all_df_list <- lapply(LPI_EU2$ID, doMean)
```

```{r, eval=FALSE}
all_matrix <- matrix(unlist(all_df_list), ncol=6, byrow=TRUE)
mean_df <- data.frame(all_matrix)
colnames(mean_df) <- c("ID", "Term","Estimate","SE","Statistic","p.val")

write.csv(mean_df, paste(wd,"/Chapter_Two/Europe/" ,"Rate_Mean_Temp_Change.csv", sep=""))

```

Land use - HILDA dataset from http://www.wageningenur.nl/en/Expertise-Services/Chair-groups/Environmental-Sciences/Laboratory-of-Geoinformation-Science-and-Remote-Sensing/Models/Hilda.htm
```{r, warning=FALSE}
library(maptools)
r_1950 <-raster(readAsciiGrid( paste(wd,"/Chapter_Two/Europe/" ,"eu27ch1950.asc", sep="")))
r_1960 <-raster(readAsciiGrid( paste(wd,"/Chapter_Two/Europe/" ,"eu27ch1960.asc", sep="")))
r_1970 <-raster(readAsciiGrid( paste(wd,"/Chapter_Two/Europe/" ,"eu27ch1970.asc", sep="")))
r_1980 <-raster(readAsciiGrid( paste(wd,"/Chapter_Two/Europe/" ,"eu27ch1980.asc", sep="")))
r_1990 <-raster(readAsciiGrid( paste(wd,"/Chapter_Two/Europe/" ,"eu27ch1990.asc", sep="")))
r_2000 <-raster(readAsciiGrid( paste(wd,"/Chapter_Two/Europe/" ,"eu27ch2000.asc", sep="")))
r_2010 <-raster(readAsciiGrid( paste(wd,"/Chapter_Two/Europe/" ,"eu27ch2010.asc", sep="")))

r_2010b<-crop(r_2010, r_1950)    #the 2010 raster is larger than the other years so has to be ropped to the same size so that they can be stacked

r<-stack(r_1950,r_1960,r_1970,r_1980,r_1990,r_2000,r_2010b)

#writeRaster(r, paste(wd,"/Chapter_Two/Europe/" ,"Land_Use_1950_2010.grd", sep=""))

names(r)<-seq(1950,2010, by=10)

crs.geo <- CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")

proj4string(r) <- crs.geo

```

```{r, warning =FALSE}
library(rgdal)

xy<-cbind(LPI_EU$Longitude, LPI_EU$Latitude)

id<-LPI_EU$ID

xy_eu<-project(xy,"+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")

check_xy<-data.frame(id,xy_eu)      #subsetting the populations so that only those within the landuse raster are included - means that non-EU27 populations are lost

check_xy$check<-extract(r,check_xy[,(2:3)])

check_xy2<-na.omit(check_xy)
#check_xy2<-subset(check_xy, check != "NA")

crop_id<-check_xy2[,1]

selectedRows <- (LPI_EU$ID %in% check_xy2$id)
LPI_EU_sel <- LPI_EU[selectedRows,]
points <- SpatialPoints(cbind(check_xy2[,2], check_xy2[,3]))

x<-check_xy2[,2]
y<-check_xy2[,3]

xy<-cbind(x,y)        #creating the grid around the population - 5km either side gives a grid of 121km^2

xmin<- colFromX(r[[1]], points[1:length(points),]) - 5
xmax<- colFromX(r[[1]], points[1:length(points),]) + 5
ymin<- rowFromY(r[[1]], points[1:length(points),]) - 5
ymax<- rowFromY(r[[1]], points[1:length(points),]) + 5    #changed from 5 to 12



grid_crop<-data.frame(ID=crop_id,xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax)   #setting extents of grid to extract

grid_crop2<-merge(grid_crop,LPI_EU_sel, by="ID")

grid_crop2<-subset(grid_crop2, Specific_location ==1) 




```

```{r}

result <- data.frame() #empty result dataframe


for (i in 1:length(grid_crop2$ID)) {
  
  ID<-grid_crop2[i,1]
  Binomial<-as.character(grid_crop2[i,7])
  spid = grid_crop2[i,70:132]                     #subsetting only the dates
  colnames(spid)<-1950:2012
  
  Date<-as.numeric(colnames(spid))
  spidt<-destring(t(spid))
  
  Year<-Date[min(which(!is.na(spidt))):max(which(!is.na(spidt)))]
  Population<-spidt[min(which(!is.na(spidt))):max(which(!is.na(spidt)))]
  
  LUmin<-min(round_any(Year, 10))
  LUmax<-max(round_any(Year, 10))
  
  crop_check<-crop(r, extent(r, grid_crop2[i,4],grid_crop2[i,5],grid_crop2[i,2],grid_crop2[i,3]))
  
  crop_df<-data.frame(as.matrix(crop_check))
  
  #crop_df<-na.omit(crop_df)
  
  colnames(crop_df)<-seq(1950,2010, by=10)
  
  col<-as.numeric(colnames(crop_df[,1:7]))
  
  min_yr<-min(which(col>=LUmin))
  max_yr<-max(which(col<=LUmax))
  
  if (min_yr != max_yr) {
    
    crop_df$check<-rowSums(crop_df[,min_yr] == crop_df[,c(min_yr:max_yr)])
    
  }else{
    crop_df$check<-1
  }

  crop_df$change<-  crop_df$check != length(seq(min_yr,max_yr, by=1))
  
  change_rate<-sum(na.omit(crop_df$change=="TRUE"))/length(na.omit(crop_df$change))
  
  #print(change_rate)
  
  final<-cbind(ID,Binomial,change_rate) 
  result<-rbind(final,result)
  
}

```


```{r, eval=FALSE}

write.csv(result, paste(wd,"/Chapter_Two/Europe/" ,"LPI_LandUse_121_18_09_2015.csv", sep="" ))
LPI_LUC<-merge(LPI_EU_sel, result[,c(1,3)], by="ID",all = TRUE)


```

```{r, warning=FALSE, message=FALSE, eval=FALSE}

bm<-read.csv("Amniote_Database_Aug_2015.csv")
bm$Binomial<-paste(bm$genus, bm$species, sep="_")

lpibm<-merge(LPI, bm, by="Binomial")

lpibm<-data.frame(lpibm$Binomial, lpibm$ID, lpibm$adult_body_mass_g)
head(lpibm)
colnames(lpibm)<-c("Binomial", "ID", "Body_mass_g")

lpibm$Log_Body_Mass_g<-log10(lpibm$Body_mass_g)

lpibm2<-lpibm[lpibm$Body_mass_g !=-999,]


lpibm2<-unique(lpibm2)

lpibm2$Log_Body_Mass_g<-log10(lpibm2$Body_mass_g)

write.csv(lpibm2, paste(wd, "/Chapter_Two/Europe/" ,"LPI_BodyMass_Amniote_Database.csv", sep=""))


```

```{r}
lpi<-read.csv("LPI_pops_20160523_edited.csv")
lpi<-lpi[,c(1,15,32,31)]
pop<-read.csv(paste(wd,"/Chapter_Two/Europe/", "Global_Population_Lambdas.csv", sep=""))
pop<-pop[,c(2,5,8,9)]
clim<-read.csv(paste(wd,"/Chapter_Two/Europe/" ,"Rate_Mean_Temp_Change.csv", sep=""))
clim<-clim[,c(2,4)]
luc<- read.csv(paste(wd,"/Chapter_Two/Europe/" ,"LPI_LandUse_121_18_09_2015.csv", sep="" ))
luc<-luc[,c(2,4)]
bm<- read.csv(paste(wd, "/Chapter_Two/Europe/", "LPI_BodyMass_Amniote_Database.csv", sep=""))
bm<-bm[,-1]

df<-merge(merge(pop, luc,by="ID", all=T), merge(clim, bm, by="ID", all=T), by="ID", all=T)
df<-merge(df, lpi, by="ID")
head(df)
```

#Mammals
```{r}
df_mammal<-subset(df, !is.na(lambda_mean)& !is.na(change_rate) & !is.na(Estimate) & !is.na(Binomial) & length_time > 5 & Class=="Mammalia")
nrow(df_mammal)

```

```{r, message=FALSE, warning=FALSE}

library(data.table)
library(plyr)

sp_dups<-data.frame(ddply(df_mammal,.(Longitude,Latitude),nrow))
sp_dups$loc_id<-1:length(sp_dups$Longitude)
df_mammal<-merge(sp_dups, df_mammal, by=c("Longitude","Latitude"))


parm_df<-df_mammal[,c("ID","change_rate", "Estimate", "Log_Body_Mass_g")]  ##ID, land use, and climate

parm_mat<-as.matrix(parm_df)
parm_scale_m<-scale(parm_mat[,c("change_rate", "Estimate", "Log_Body_Mass_g")])       #use the scaling factors at the bottom of these to scale the rasters

parm_id<-parm_mat[,"ID"]

parm_df_scale<-data.frame(parm_id,parm_scale_m)

colnames(parm_df_scale)<-c("ID","change_rate_scale","mean_slope_scale", "body_mass_scale" )

sp_df_scale<-merge(df_mammal, parm_df_scale, by="ID")

dt_mammal<-data.table(sp_df_scale)

```

```{r, message=FALSE, warning=FALSE}

source("rsquaredglmm.R")

  library(lme4) 
  library(MuMIn)
  
  mm0<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+change_rate_scale:mean_slope_scale+body_mass_scale+(1|Binomial)+(1|loc_id),data=dt_mammal, REML=F)
  mm0a<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+body_mass_scale+(1|Binomial)+(1|loc_id),data=dt_mammal, REML=F)
  mm0b<-lmer(lambda_mean ~ change_rate_scale+body_mass_scale+(1|Binomial)+(1|loc_id),data=dt_mammal, REML=F)
  mm0c<-lmer(lambda_mean ~ mean_slope_scale+body_mass_scale+(1|Binomial)+(1|loc_id),data=dt_mammal, REML=F)
  mm0d<-lmer(lambda_mean ~ body_mass_scale+(1|Binomial)+(1|loc_id),data=dt_mammal, REML=F)
  mm1<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+change_rate_scale:mean_slope_scale+(1|Binomial)+(1|loc_id),data=dt_mammal, REML=F)
  mm1a<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+(1|Binomial)+(1|loc_id),data=dt_mammal, REML=F)
  mm1b<-lmer(lambda_mean ~ change_rate_scale+(1|Binomial)+(1|loc_id),data=dt_mammal, REML=F)
  mm1c<-lmer(lambda_mean ~ mean_slope_scale+(1|Binomial)+(1|loc_id),data=dt_mammal, REML=F)
  mmnull<-lmer(lambda_mean ~ 1+(1|Binomial)+(1|loc_id),data=dt_mammal, REML=F)
  mmsAICc <- model.sel(mm0,mm0a,mm0b,mm0c,mm0d,mm1,mm1a,mm1b,mm1c,mmnull)
  mmsAICc$model<-rownames(mmsAICc)
  mmsAICc<-data.frame(mmsAICc)
  mmsAICc
  
  AIC(mm0,mm0a,mm0b,mm0c,mm0d,mm1,mm1a,mm1b,mm1c,mmnull)

  mmodels_list<-list(mm0,mm0a,mm0b,mm0c,mm0d,mm1,mm1a,mm1b,mm1c,mmnull)
  mmodelsR<-lapply(mmodels_list,rsquared.glmm)
  mmodelsRsq <- matrix(unlist(mmodelsR), ncol=6, byrow=T)
  rownames(mmodelsRsq)<-c("m0","m0a","m0b","m0c","m0d","m1","m1a","m1b","m1c","mnull")

  mmodelsRsq
```

```{r}
  library(MuMIn)
  mvar_imp<-summary(model.avg(mmodels_list))

  #mav<-model.avg(models_list, subset = cumsum(weight) <= 0.95)
  mmav<-model.avg(mmodels_list, subset = delta <= 6)
  
  msmav<-summary(mmav)
  
  mcoef_av<-msmav$coefmat.subset[1:5,"Estimate"]
  mcoef_df<-data.frame(mcoef_av)
  mcoef_df$lowCI<-confint(mmav)[1:5,1]
  mcoef_df$highCI<-confint(mmav)[1:5,2]
  mcoef_df
  
  
  mcoef_pcnt<-data.frame(((10^mcoef_df) - 1)*100)
  mcoef_pcnt

``` 

#Birds
```{r}
df_birds<-subset(df, !is.na(lambda_mean)& !is.na(change_rate) & !is.na(Estimate) & !is.na(Binomial) & length_time > 5 & Class=="Aves")
nrow(df_birds)

```

```{r, message=FALSE, warning=FALSE}

library(data.table)

sp_dups<-data.frame(ddply(df_birds,.(Longitude,Latitude),nrow))
sp_dups$loc_id<-1:length(sp_dups$Longitude)
df_birds<-merge(sp_dups, df_birds, by=c("Longitude","Latitude"))


parm_df<-df_birds[,c("ID","change_rate", "Estimate", "Log_Body_Mass_g")]  ##ID, land use, and climate

parm_mat<-as.matrix(parm_df)
parm_scale_b<-scale(parm_mat[,c("change_rate", "Estimate", "Log_Body_Mass_g")])       #use the scaling factors at the bottom of these to scale the rasters

parm_id<-parm_mat[,"ID"]

parm_df_scale<-data.frame(parm_id,parm_scale_b)

colnames(parm_df_scale)<-c("ID","change_rate_scale","mean_slope_scale", "body_mass_scale" )

sp_df_scale<-merge(df_birds, parm_df_scale, by="ID")

dt_birds<-data.table(sp_df_scale)

```

```{r, message=FALSE, warning=FALSE}

source("rsquaredglmm.R")

  library(lme4) 
  library(MuMIn)
  
  bm0<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+change_rate_scale:mean_slope_scale+body_mass_scale+(1|Binomial)+(1|loc_id),data=dt_birds, REML=F)
  bm0a<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+body_mass_scale+(1|Binomial)+(1|loc_id),data=dt_birds, REML=F)
  bm0b<-lmer(lambda_mean ~ change_rate_scale+body_mass_scale+(1|Binomial)+(1|loc_id),data=dt_birds, REML=F)
  bm0c<-lmer(lambda_mean ~ mean_slope_scale+body_mass_scale+(1|Binomial)+(1|loc_id),data=dt_birds, REML=F)
  bm0d<-lmer(lambda_mean ~ body_mass_scale+(1|Binomial)+(1|loc_id),data=dt_birds, REML=F)
  bm1<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+change_rate_scale:mean_slope_scale+(1|Binomial)+(1|loc_id),data=dt_birds, REML=F)
  bm1a<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+(1|Binomial)+(1|loc_id),data=dt_birds, REML=F)
  bm1b<-lmer(lambda_mean ~ change_rate_scale+(1|Binomial)+(1|loc_id),data=dt_birds, REML=F)
  bm1c<-lmer(lambda_mean ~ mean_slope_scale+(1|Binomial)+(1|loc_id),data=dt_birds, REML=F)
  bmnull<-lmer(lambda_mean ~ 1+(1|Binomial)+(1|loc_id),data=dt_birds, REML=F)
  bmsAICc <- model.sel(bm0,bm0a,bm0b,bm0c,bm0d,bm1,bm1a,bm1b,bm1c,bmnull)
  bmsAICc$model<-rownames(bmsAICc)
  bmsAICc<-data.frame(bmsAICc)
  bmsAICc
  
  AIC(bm0,bm0a,bm0b,bm0c,bm0d,bm1,bm1a,bm1b,bm1c,bmnull)

  bmodels_list<-list(bm0,bm0a,bm0b,bm0c,bm0d,bm1,bm1a,bm1b,bm1c,bmnull)
  bmodelsR<-lapply(bmodels_list,rsquared.glmm)
  bmodelsRsq <- matrix(unlist(bmodelsR), ncol=6, byrow=T)
  rownames(bmodelsRsq)<-c("m0","m0a","m0b","m0c","m0d","m1","m1a","m1b","m1c","mnull")

  bmodelsRsq
    


```

```{r}
  library(MuMIn)
  bvar_imp<-summary(model.avg(bmodels_list))

  #mav<-model.avg(models_list, subset = cumsum(weight) <= 0.95)
  bmav<-model.avg(bmodels_list, subset = delta <= 6)
  
  bsmav<-summary(bmav)
  
  nums<-c("1", "2", "3", "12", "13", "23", "123", "234", "1234", "(Null)")
  mods<-c("bm0d", "bm1b", "bm1c", "bm0b", "bm0c", "bm1a", "bm0a", "bm1", "bm0", "bmnull")
  mod_num<-cbind(nums, mods)
  b_models<-rownames(bvar_imp$msTable)
  b_weights<-Weights(bmav)
  b_mod_weights<-cbind(b_models, b_weights)
  b_model_weights<-merge(b_mod_weights, mod_num, by.x="b_models", by.y="nums")  
  
  bcoef_av<-bsmav$coefmat.subset[1:5,"Estimate"]
  bcoef_df<-data.frame(bcoef_av)
  bcoef_df$lowCI<-confint(bmav)[1:5,1]
  bcoef_df$highCI<-confint(bmav)[1:5,2]
  bcoef_df
  
  
  bcoef_pcnt<-data.frame(((10^bcoef_df) - 1)*100)
  bcoef_pcnt

```  
    
    
    
```{r, warning=FALSE}
bcoef_pcnt$Class<-"Birds"  
mcoef_pcnt$Class<-"Mammals"  

bcoef_pcnt$Var<-rownames(bcoef_pcnt)
mcoef_pcnt$Var<-rownames(mcoef_pcnt)
colnames(bcoef_pcnt)[1]<-"ave"
colnames(mcoef_pcnt)[1]<-"ave"

coef_pcnt<-rbind(bcoef_pcnt, mcoef_pcnt)

library(ggplot2)
p1<-ggplot(coef_pcnt, aes(colour=Class))
p1<- p1 + geom_hline(yintercept = 0, colour=gray(1/2), lty=2)
p1<- p1 + geom_linerange(aes(x=Var, ymin=lowCI, ymax=highCI), lwd=2.5, position = position_dodge(width=2/3))
p1<- p1 + geom_pointrange(aes(x= Var, y=ave, ymin=lowCI, ymax=highCI), lwd=2, position=position_dodge(width=2/3), shape=21, fill="White")
#p1<- p1 + scale_y_continuous(breaks=seq(-8, 6, 2), limits=(c(-9,5))) +theme_bw() + labs(y = "Population Change (%)", x = "") + theme(legend.position="none",text=element_text(size=20),axis.text.x=element_text(size=8) , axis.title.x = element_text(margin = unit(c(5, 0, 0, 0), "mm")))
#p1<- p1 + scale_color_manual(values=c("black", "black"))
p1<-p1+coord_flip()
print(p1)

  

```


####Spatial Predictions - Europe####

```{r Prediction maps - climate}
mean_t<- brick(paste(wd,"/Chapter_Two/Europe/" ,"Europe_mean_mon.grd", sep=""))
mean_t2<-trim(mean_t)

x<-seq(-38.625, 75.125, by=0.25)
y<-seq(25.375, 71.875, by=0.25)

xy<-expand.grid(x,y)
colnames(xy)<-c("Lon", "Lat")


time<-1:nlayers(mean_t)
## add 1 for a model with an intercept
X <- cbind(1, time)

## pre-computing constant part of least squares
invXtX <- solve(t(X) %*% X) %*% t(X)

## much reduced regression model; [2] is to get the slope
quickfun <- function(y) (invXtX %*% y)[2]
x4 <- calc(mean_t, quickfun)

```

```{r, eval=FALSE}
writeRaster(x4, filename=paste(wd,"/Chapter_Two/Europe/" ,"climate_change_slope.tif", sep=""), format="GTiff", overwrite=TRUE)

```

```{r, eval=FALSE}
r<-brick(paste(wd,"/Chapter_Two/Europe/" ,"Land_Use_1950_2010.grd", sep=""))

crs.geo <- CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")

proj4string(r) <- crs.geo

names(r)<-seq(1950,2010, by=10)
plot(r)

r_df<-data.frame(as.matrix(r))

r_df$a<-as.numeric(r_df[,1] != r_df[,2])   #checking for each decade if there is a change in land use category
r_df$b<-as.numeric(r_df[,2] != r_df[,3])
r_df$c<-as.numeric(r_df[,3] != r_df[,4])
r_df$d<-as.numeric(r_df[,4] != r_df[,5])
r_df$e<-as.numeric(r_df[,5] != r_df[,6])
r_df$f<-as.numeric(r_df[,6] != r_df[,7])

r_df$total_change<-rowSums(r_df[,c(8:13)])   #total number of category changes - a way of quantifying how unstable a landscape is?


r_rast<-raster(matrix(r_df$total_change, nrow=4075,ncol=4011, byrow=T))

plot(r_rast)

extent(r_rast)<-c(2554419, 6565419, 1260409, 5335409)
crs(r_rast)<-"+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs"

r_rast

n <- c(0,0.99,0,1,6,1)
m <- matrix(n, ncol=3, byrow=TRUE)

rc <- reclassify(r_rast, m)     #reclassifying to binary rather than number of changes

f<-rep.int(1, 121)
m<-matrix(f, nrow=11)

my_fun2 <- function(x) {
  result <- sum(na.omit(x))/length(na.omit(x))
  return(result)
}


rm <- focal(rc, w=matrix(1,nrow=11,ncol=11), fun=my_fun2)    #looking at amount of change in the surrounding area

plot(rm)



```

```{r, eval=FALSE}
writeRaster(rm, paste(wd,"/Chapter_Two/Europe/" ,"land_use_change.tif", sep=""), format="GTiff", overwrite=TRUE )


```


Spatial Prediction - Europe Birds

```{r, warning=FALSE}

library(raster)
LU<-raster( paste(wd,"/Chapter_Two/Europe/" ,"land_use_change.tif", sep=""))
TM<-raster(paste(wd,"/Chapter_Two/Europe/" ,"climate_change_slope.tif", sep=""))

crs_laea<-"+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs"
extent(LU)<-c(2554419, 6565419, 1260409, 5335409)
crs(LU)<-crs_laea

TM<-projectRaster(TM, crs = crs_laea)
#plot(TM)

centre_temp_b<-attr(parm_scale_b, 'scaled:center')[2]
centre_luc_b<-attr(parm_scale_b, 'scaled:center')[1]
scale_temp_b<-attr(parm_scale_b, 'scaled:scale')[2] 
scale_luc_b<-attr(parm_scale_b, 'scaled:scale')[1] 

LU_b<-(LU  - centre_luc_b )/scale_luc_b
TM_b<-(TM -centre_temp_b)/ scale_temp_b
LU_bt<-trim(LU_b)
TM_b2<-resample(TM_b, LU_bt)
#LU_bt2<-resample(LU_ust,TM_us)

#bm<-reclassify(TM_b, c(-1,max(na.omit(values(TM_b))),0))
bird_stack<-stack(TM_b2,LU_bt)

names(bird_stack)<-c("mean_slope_scale","change_rate_scale")

# bird_average<-function(model, weight){
#   pred<-predict(bird_stack,model, re.form=NA)
#   pred_weight<-pred*weight
#   return(pred_weight)
#   }
# 
# b_mod_weights$b_weights<-as.numeric(as.character(b_model_weights$b_weights))
# 
# mapply(bird_average, model=bmodels_list, weight=b_model_weights$b_weights)

pred<-predict(bird_stack,bm1, re.form=NA)

plot(trim(10^pred))

xy <- data.frame(lon=dt_birds$Longitude, lat=dt_birds$Latitude)
coordinates(xy) <- c("lon", "lat")
proj4string(xy) <- CRS("+init=epsg:4326") # WGS 84 UTM 35S
CRS.new <- CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")
xy_new<- spTransform(xy, CRS.new)
points(xy_new)

```


Spatial Prediction - Europe Mammals
```{r}

library(raster)
LU<-raster( paste(wd,"/Chapter_Two/Europe/" ,"land_use_change.tif", sep=""))
TM<-raster(paste(wd,"/Chapter_Two/Europe/" ,"climate_change_slope.tif", sep=""))

crs_laea<-"+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs"
extent(LU)<-c(2554419, 6565419, 1260409, 5335409)
crs(LU)<-crs_laea

TM<-projectRaster(TM, crs = crs_laea)
#plot(TM)

centre_temp_m<-attr(parm_scale_m, 'scaled:center')[2]
centre_luc_m<-attr(parm_scale_m, 'scaled:center')[1]
scale_temp_m<-attr(parm_scale_m, 'scaled:scale')[2] 
scale_luc_m<-attr(parm_scale_m, 'scaled:scale')[1] 

LU_m<-(LU  - centre_luc_m )/scale_luc_m
TM_m<-(TM -centre_temp_m)/ scale_temp_m
LU_mt<-trim(LU_m)
TM_m2<-resample(TM_m, LU_mt)

mammal_stack<-stack(TM_m2,LU_mt)

names(mammal_stack)<-c("mean_slope_scale","change_rate_scale")


pred<-predict(mammal_stack,mm1, re.form=NA)

plot(trim(10^pred))

xy <- data.frame(lon=dt_mammal$Longitude, lat=dt_mammal$Latitude)
coordinates(xy) <- c("lon", "lat")
proj4string(xy) <- CRS("+init=epsg:4326") # WGS 84 UTM 35S
CRS.new <- CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")
xy_new<- spTransform(xy, CRS.new)
points(xy_new)
```



####Harmonized LUH dataset - Global / CRU Temps Global Climate####

```{r}



```

#ESA dataset - Global

```{r}



```

#Spatial Current and Future Predictions - Global
```{r, warning=FALSE}

df_bird<-read.csv("bird_data_for_prediction.csv")
df_mammal<-read.csv("mammal_data_for_prediction.csv")

# df_bird[is.na(df_bird$lambda_mean),]$lambda_mean<-0
# df_mammal[is.na(df_mammal$lambda_mean),]$lambda_mean<-0

library(plyr)
#counting duplicates at each location
sp_dups_bird<-data.frame(ddply(df_bird,.(Longitude,Latitude),nrow))
sp_dups_mamm<-data.frame(ddply(df_mammal,.(Longitude,Latitude),nrow))

sp_dups_bird$loc_id<-1:length(sp_dups_bird$Longitude)
sp_dups_mamm$loc_id<-1:length(sp_dups_mamm$Longitude)

sp_dups_df_bird<-merge(sp_dups_bird, df_bird, by=c("Longitude","Latitude"))
sp_dups_df_mamm<-merge(sp_dups_mamm, df_mammal, by=c("Longitude","Latitude"))

library(data.table)
dt_bird = as.data.table(sp_dups_df_bird)
dt_mamm = as.data.table(sp_dups_df_mamm)

parm_df_bird<-sp_dups_df_bird[,c("ID","Estimate", "both_change", "Log_Body_Mass_g")]  ##ID, land use, and climate  use "LUC_dist" or "Nat_change" for purely annual change in summed primary, secondary and other
parm_df_mamm<-sp_dups_df_mamm[,c("ID","Estimate", "both_change", "Log_Body_Mass_g")]  ##ID, land use, and climate  use "LUC_dist" or "Nat_change" for purely annual change in summed primary, secondary and othe

parm_mat_bird<-as.matrix(parm_df_bird)
parm_mat_mamm<-as.matrix(parm_df_mamm)

parm_scale_bird<-scale(parm_mat_bird[,c("Estimate", "both_change", "Log_Body_Mass_g")])       #use the scaling factors at the bottom of these to scale the rasters
parm_scale_mamm<-scale(parm_mat_mamm[,c("Estimate", "both_change", "Log_Body_Mass_g")])       #use the scaling factors at the bottom of these to scale the rasters

parm_id_bird<-parm_mat_bird[,"ID"]
parm_id_mamm<-parm_mat_mamm[,"ID"]

parm_df_scale_bird<-data.frame(parm_id_bird,parm_scale_bird)
parm_df_scale_mamm<-data.frame(parm_id_mamm,parm_scale_mamm)

colnames(parm_df_scale_bird)<-c("ID","mean_slope_scale", "change_rate_scale", "Bodymass_scale")
colnames(parm_df_scale_mamm)<-c("ID","mean_slope_scale", "change_rate_scale", "Bodymass_scale")

sp_df_scale_bird<-merge(sp_dups_df_bird, parm_df_scale_bird, by="ID")
sp_df_scale_mamm<-merge(sp_dups_df_mamm, parm_df_scale_mamm, by="ID")

dt_bird<-data.table(sp_df_scale_bird)
dt_mamm<-data.table(sp_df_scale_mamm)
```


```{r}
library(lme4)
m1c_b<-lmer(lambda_mean ~ mean_slope_scale+(1|Binomial)+(1|loc_id),data=dt_bird, REML=F)
m1c_m<-lmer(lambda_mean ~ mean_slope_scale+(1|Binomial)+(1|loc_id),data=dt_mamm, REML=F)

m1_m<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+change_rate_scale:mean_slope_scale+(1|Binomial)+(1|loc_id),data=dt_mamm, REML=F)
```

```{r, cache=TRUE}
###############################################
library(raster)

CR40s<-brick("cru_ts3.23.1941.1950.tmp.dat.nc")
CR50s<-brick("cru_ts3.23.1951.1960.tmp.dat.nc")
CR60s<-brick("cru_ts3.23.1961.1970.tmp.dat.nc")
CR70s<-brick("cru_ts3.23.1971.1980.tmp.dat.nc")
CR80s<-brick("cru_ts3.23.1981.1990.tmp.dat.nc")
CR90s<-brick("cru_ts3.23.1991.2000.tmp.dat.nc")
CR00s<-brick("cru_ts3.23.2001.2010.tmp.dat.nc")

obs<-stack(CR40s[[109:120]], CR50s, CR60s, CR70s, CR80s, CR90s,CR00s[[1:48]] )
year_mon<-rep(1:(nlayers(obs)/12), each=12)


#changing from monthly to annual data
year_s<-stackApply(obs, indices=year_mon, fun=mean, na.rm=TRUE)

years<-1950:2004
names(year_s)<-years

X <- cbind(1, years)

## pre-computing constant part of least squares
invXtX <- solve(t(X) %*% X) %*% t(X)

## much reduced regression model; [2] is to get the slope
quickfun <- function(y) (invXtX %*% y)[2]

x4 <- calc(year_s, quickfun)

plot(x4)


```

```{r}
#####HadGEM ES

### rcp 85

ts85_05_30<-brick("ts_Amon_HadGEM2-ES_esmrcp85_r1i1p1_200512-203011.nc")
ts85_30_55<-brick("ts_Amon_HadGEM2-ES_esmrcp85_r1i1p1_203012-205511.nc")
ts85_55_80<-brick("ts_Amon_HadGEM2-ES_esmrcp85_r1i1p1_205512-208011.nc")
ts85_80_00<-brick("ts_Amon_HadGEM2-ES_esmrcp85_r1i1p1_208012-210011.nc")

ts85_05_00<-stack(ts85_05_30,ts85_30_55,ts85_55_80,ts85_80_00)

### rcp 26

ts26_05_30<-brick("ts_Amon_HadGEM2-ES_rcp26_r1i1p1_200512-203011.nc")
ts26_30_55<-brick("ts_Amon_HadGEM2-ES_rcp26_r1i1p1_203012-205511.nc")
ts26_55_80<-brick("ts_Amon_HadGEM2-ES_rcp26_r1i1p1_205512-208011.nc")
ts26_80_00<-brick("ts_Amon_HadGEM2-ES_rcp26_r1i1p1_208012-209911.nc")
ts26_00_12<-brick("ts_Amon_HadGEM2-ES_rcp26_r1i1p1_209912-212411.nc")

ts26_05_00<-stack(ts26_05_30,ts26_30_55,ts26_55_80,ts26_80_00, ts26_00_12[[1:12]])
#ts_05_00<-brick(ts_05_00)

```

```{r}
#rcp85
modelcc<-rotate(ts85_05_00)

obs_r<-suppressWarnings(resample(obs, modelcc))

obs_rk<-obs_r+273.15

mask_mod<-mask(modelcc, obs_rk[[1]])

#obs_mod<-stack(obs_rk, modelcc)

obs_mod<-stack(obs_rk, mask_mod)

obs_mod_mon_mean<-matrix(cellStats(obs_mod, mean))

plot(obs_mod_mon_mean)

#############################################

```


```{r}
#rcp26
modelcc_26<-rotate(ts26_05_00)

obs_r_26<-suppressWarnings(resample(obs, modelcc_26))


obs_rk_26<-obs_r_26+273.15

mask_mod_26<-mask(modelcc_26, obs_rk_26[[1]])

#obs_mod<-stack(obs_rk, modelcc)

obs_mod_26<-stack(obs_rk_26, mask_mod_26)

obs_mod_mon_mean_26<-matrix(cellStats(obs_mod_26, mean))

plot(obs_mod_mon_mean_26)

#############################################

```



```{r}
#rcp85
year_mon<-rep(1:(nlayers(obs_mod)/12), each=12)

#changing from monthly to annual data
obs_mod_ann<-stackApply(obs_mod, indices=year_mon, fun=mean, na.rm=TRUE)

obs_mat_ann<-matrix(cellStats(obs_mod_ann, mean))

plot(obs_mat_ann)


```

```{r}
#rcp26
year_mon_26<-rep(1:(nlayers(obs_mod_26)/12), each=12)

#changing from monthly to annual data
obs_mod_ann_26<-stackApply(obs_mod_26, indices=year_mon_26, fun=mean, na.rm=TRUE)

obs_mat_ann_26<-matrix(cellStats(obs_mod_ann_26, mean))
years<-1951:2100
plot(years,obs_mat_ann_26 - 273.15 )

df<-data.frame(years, obs_mat_ann_26 - 273.15)
colnames(df)[2]<-"Mean_Temp"
df$Observed_Predict<-as.factor(c(rep("Observed",54), rep("Predicted",96)))

ggplot(df, aes(x= years, y= Mean_Temp, colour = Observed_Predict))+
  geom_point() + 
  ylim(7.5, 17) +
  labs(y = expression(Global~Mean~Temperature^{o}~C), x = "Year", colour = "", title = "RCP 2.6")
  
  
```


```{r}
#rcp85
x = c(1:150)
g = mgcv:::gam(obs_mat_ann~s(x, k=8), fx=TRUE)
plot(g)

g_pred = predict(g)
g_pred = g_pred[55:150]
d = diff(g_pred)
plot(g_pred)
plot(d)
```


```{r}
#rcp26
x = c(1:150)
g_26 = mgcv:::gam(obs_mat_ann_26~s(x, k=8), fx=TRUE)
plot(g_26)

g_pred_26 = predict(g_26)
g_pred_26 = g_pred_26[55:150]
d_26 = diff(g_pred_26)
plot(g_pred_26)
plot(d_26)
```


```{r}
#bird values
centre_tempb<-0.045237717 
scale_tempb<-0.071445065 

#mammal values
centre_tempm<-0.015412738 
scale_tempm<-0.079118974
```


```{r}
#rcp85
gam_diff_b<-scale(d, center=centre_tempb, scale=scale_tempb)
plot(gam_diff_b)

gam_diff_m<-scale(d, center=centre_tempm, scale=scale_tempm)
plot(gam_diff_m)
```

```{r}
#rcp26
gam_diff_b_26<-scale(d_26, center=centre_tempb, scale=scale_tempb)
plot(gam_diff_b_26)

gam_diff_m_26<-scale(d_26, center=centre_tempm, scale=scale_tempm)
plot(gam_diff_m_26)
```



```{r}
#rcp85
ave_mod_b<-data.frame(gam_diff_b, rep(0, length(gam_diff_b)), rep(0, length(gam_diff_b)))
colnames(ave_mod_b)<-c("mean_slope_scale", "change_rate_scale", "Bodymass_scale")
pop_pred_b<-stats:::predict(m1c_b, ave_mod_b, re.form=NA, se.fit=TRUE)

ave_mod_m<-data.frame(gam_diff_m, rep(0, length(gam_diff_m)), rep(0, length(gam_diff_m)))
colnames(ave_mod_m)<-c("mean_slope_scale", "change_rate_scale", "Bodymass_scale")
pop_pred_m<-predict(m1_m, ave_mod_m, re.form=NA, se.fit=T)
```

```{r, warning=FALSE, message=FALSE}
#rcp26
ave_mod_b_26<-data.frame(gam_diff_b_26, rep(0, length(gam_diff_b_26)), rep(0, length(gam_diff_b_26)))
colnames(ave_mod_b_26)<-c("mean_slope_scale", "change_rate_scale", "Bodymass_scale")
pop_pred_b_26<-predict(m1c_b, ave_mod_b_26, re.form=NA, se.fit=T)

ave_mod_m_26<-data.frame(gam_diff_m_26, rep(0, length(gam_diff_m_26)), rep(0, length(gam_diff_m_26)))
colnames(ave_mod_m_26)<-c("mean_slope_scale", "change_rate_scale", "Bodymass_scale")
pop_pred_m_26<-stats:::predict(m1_m, ave_mod_m_26, re.form=NA, se.fit=T)
```


```{r}
#rcp85
plot(pop_pred_b$fit, ylim=c(-0.045, 0.005), main="Birds RCP 8.5", type="l")
lines(pop_pred_b$fit + 1.96*pop_pred_b$se.fit, col="red", lty=2)
lines(pop_pred_b$fit - 1.96*pop_pred_b$se.fit, col="red", lty=2)

plot(pop_pred_m$fit, ylim=c(-0.02, 0.01), main="Mammals RCP 8.5",  type="l")
lines(pop_pred_m$fit + 1.96*pop_pred_m$se.fit, col="red", lty=2)
lines(pop_pred_m$fit - 1.96*pop_pred_m$se.fit, col="red", lty=2)
```

```{r}
#rcp26
plot(pop_pred_b_26$fit, ylim=c(-0.045, 0.025), main="Birds RCP 2.6", type="l")
lines(pop_pred_b_26$fit + 1.96*pop_pred_b_26$se.fit, col="red", lty=2)
lines(pop_pred_b_26$fit - 1.96*pop_pred_b_26$se.fit, col="red", lty=2)

plot(pop_pred_m_26$fit, ylim=c(-0.015, 0.015), main="Mammals RCP 2.6",  type="l")
lines(pop_pred_m_26$fit + 1.96*pop_pred_m_26$se.fit, col="red", lty=2)
lines(pop_pred_m_26$fit - 1.96*pop_pred_m_26$se.fit, col="red", lty=2)
```


```{r}
10^sum(pop_pred_b$fit + 1.96*pop_pred_b$se.fit)
10^sum(pop_pred_b$fit - 1.96*pop_pred_b$se.fit)
10^sum(pop_pred_b$fit)

10^sum(pop_pred_m$fit + 1.96*pop_pred_m$se.fit)
10^sum(pop_pred_m$fit - 1.96*pop_pred_m$se.fit)
10^sum(pop_pred_m$fit)
```

```{r}
#rcp85
pop_pred_b$fit2<-c(0, pop_pred_b$fit)
pop_pred_b$se.fit2<-c(0, pop_pred_b$se.fit)

pop_pred_m$fit2<-c(0, pop_pred_m$fit)
pop_pred_m$se.fit2<-c(0, pop_pred_m$se.fit)
```

```{r}
#rcp26
pop_pred_b_26$fit2<-c(0, pop_pred_b_26$fit)
pop_pred_b_26$se.fit2<-c(0, pop_pred_b_26$se.fit)

pop_pred_m_26$fit2<-c(0, pop_pred_m_26$fit)
pop_pred_m_26$se.fit2<-c(0, pop_pred_m_26$se.fit)
```

```{r}
plot(y=10^cumsum(pop_pred_b$fit2), x=2005:2100, ylim=c(0, 1.5), type="l", ylab="Bird Population Index", xlab="Year", main="Predicted Bird Population Index Under Scenario RCP 8.5")
lines(y=10^cumsum(pop_pred_b$fit2 - 1.96*pop_pred_b$se.fit2), x=2005:2100, lty=2, col="Red")
lines(y=10^cumsum(pop_pred_b$fit2 + 1.96*pop_pred_b$se.fit2), x=2005:2100, lty=2, col="Red")
```

```{r}
plot(y=10^cumsum(pop_pred_m$fit2), x=2005:2100, ylim=c(0, 3), type="l", ylab="Mammal Population Index", xlab="Year", main="Predicted Mammal Population Trends Under RCP 8.5")
lines(y=10^cumsum(pop_pred_m$fit2 - 1.96*pop_pred_m$se.fit2), x=2005:2100, lty=2, col="Red")
lines(y=10^cumsum(pop_pred_m$fit2 + 1.96*pop_pred_m$se.fit2), x=2005:2100, lty=2, col="Red")

```

```{r}
#Birds RCP 8.5
yb<-10^cumsum(pop_pred_b$fit2)
xb<-2005:2100
lcib<-10^cumsum(pop_pred_b$fit2 - 1.96*pop_pred_b$se.fit2)
ucib<-10^cumsum(pop_pred_b$fit2 + 1.96*pop_pred_b$se.fit2)

dfb<-data.frame(xb,yb, lcib, ucib)

library(ggplot2)

ggplot(dfb, aes(xb,yb))+
  geom_line(size=2)+
  geom_ribbon(aes(ymin=lcib, ymax=ucib), alpha=0.3, colour=NA)+
  labs(y = "Bird Population Index RCP 8.5", x = "")+
  scale_y_continuous(breaks=seq(0, 1, 0.25))+
  theme_bw()+
  theme(text = element_text(size=16), axis.title.y = element_text(margin = unit(c(0, 5, 0, 0), "mm")))
```

```{r}
#Birds RCP 2.6
yb_26<-10^cumsum(pop_pred_b_26$fit2)
xb_26<-2005:2100
lcib_26<-10^cumsum(pop_pred_b_26$fit2 - 1.96*pop_pred_b_26$se.fit2)
ucib_26<-10^cumsum(pop_pred_b_26$fit2 + 1.96*pop_pred_b_26$se.fit2)

dfb_26<-data.frame(xb_26,yb_26, lcib_26, ucib_26)

library(ggplot2)

ggplot(dfb_26, aes(xb_26,yb_26))+
  geom_line(size=2)+
  geom_ribbon(aes(ymin=lcib_26, ymax=ucib_26), alpha=0.3, colour=NA)+
  labs(y = "Bird Population Index RCP 2.6", x = "")+
  #scale_y_continuous(breaks=seq(0, 1, 0.25))+
  theme_bw()+
  theme(text = element_text(size=16), axis.title.y = element_text(margin = unit(c(0, 5, 0, 0), "mm")))
```

```{r}
#Mammals RCP 8.5

ym<-10^cumsum(pop_pred_m$fit2)
xm<-2005:2100
lcim<-10^cumsum(pop_pred_m$fit2 - 1.96*pop_pred_m$se.fit2)
ucim<-10^cumsum(pop_pred_m$fit2 + 1.96*pop_pred_m$se.fit2)

dfm<-data.frame(xm,ym, lcim, ucim)

library(ggplot2)

ggplot(dfm, aes(xm,ym))+
  geom_line(size=2)+
  geom_ribbon(aes(ymin=lcim, ymax=ucim), alpha=0.3, colour=NA)+
  labs(y = "Mammal Population Index RCP 8.5", x = "")+
  #scale_y_continuous(breaks=seq(0, 1, 0.5))+
  theme_bw()+
  theme(text = element_text(size=16), axis.title.y = element_text(margin = unit(c(0, 5, 0, 0), "mm")))
```

```{r}
#Mammals RCP 2.6

ym_26<-10^cumsum(pop_pred_m_26$fit2)
xm_26<-2005:2100
lcim_26<-10^cumsum(pop_pred_m_26$fit2 - 1.96*pop_pred_m_26$se.fit2)
ucim_26<-10^cumsum(pop_pred_m_26$fit2 + 1.96*pop_pred_m_26$se.fit2)

dfm_26<-data.frame(xm_26,ym_26, lcim_26, ucim_26)

library(ggplot2)

ggplot(dfm_26, aes(xm_26,ym_26))+
  geom_line(size=2)+
  geom_ribbon(aes(ymin=lcim_26, ymax=ucim_26), alpha=0.3, colour=NA)+
  labs(y = "Mammal Population Index RCP 2.6", x = "")+
  #scale_y_continuous(breaks=seq(0, 1, 0.25))+
  theme_bw()+
  theme(text = element_text(size=16), axis.title.y = element_text(margin = unit(c(0, 5, 0, 0), "mm")))
```


#Spatial prediction on observed trends
```{r}
################function for doing it all spatially
#birds observed
x = c(1:nlayers(year_s))

funob<-function(y){
  
  z<-matrix(y)
  if (sum(is.na(z))<1){
    g<-mgcv:::gam(matrix(z) ~ s(x, k=8), fx=TRUE)
    g_pred<-predict(g)
    #g_pred<-g_pred[55:148]
    #g_pred<-g_pred[55:100] #2050
    d<-diff(g_pred)
    gam_diff<-scale(d, center = centre_tempb, scale = scale_tempb)
    ave_mod<-data.frame(gam_diff, rep(0, length(gam_diff)), rep(0, length(gam_diff)))
    colnames(ave_mod)<-c("mean_slope_scale", "change_rate_scale", "Bodymass_scale")
    pop_pred<-predict(m1c_b, ave_mod, re.form=NA)
  } else{
    
    #pop_pred$fit<-rep(NA,45) #2050
    pop_pred<-rep(NA,nlayers(year_s))
  }
  return(pop_pred)
}

obs_bird<-calc(year_s, funob)

```


#Future Spatial Predictions


```{r}
################function for doing it all spatially
#mammals 
x = c(1:150)

funm<-function(y){
  
  z<-matrix(y)
  if (sum(is.na(z))<1){
    g<-mgcv:::gam(matrix(z) ~ s(x, k=8), fx=TRUE)
    g_pred<-predict(g)
    g_pred<-g_pred[55:148]
    #g_pred<-g_pred[55:100] #2050
    d<-diff(g_pred)
    gam_diff<-scale(d, center = centre_tempm, scale = scale_tempm)
    ave_mod<-data.frame(gam_diff, rep(0, length(gam_diff)), rep(0, length(gam_diff)))
    colnames(ave_mod)<-c("mean_slope_scale", "change_rate_scale", "Bodymass_scale")
    pop_pred<-predict(m1_m, ave_mod, re.form=NA, se.fit=T)
  } else{
    
    #pop_pred$fit<-rep(NA,45) #2050
    pop_pred$fit<-rep(NA,93)
  }
  return(pop_pred$fit)
}

#gam_pred_rast_2050<-calc(obs_mod_ann, fun)
#gam_pred_rast<-calc(obs_mod_ann, fun)

```

```{r}
#rcp 8.5
pred_rast_mammal<-calc(obs_mod_ann, funm)

ave_lam_mammal<-calc(pred_rast_mammal, fun=mean)
plot(10^ave_lam_mammal)

pgr_mammal<-10^pred_rast_mammal
prod_lam_mammal<-calc(pgr_mammal, fun=prod)
plot(prod_lam_mammal)
```

```{r}
#rcp 2.6
pred_rast_mammal_26<-calc(obs_mod_ann_26, funm)

ave_lam_mammal_26<-calc(pred_rast_mammal_26, fun=mean)
plot(10^ave_lam_mammal_26)

pgr_mammal_26<-10^pred_rast_mammal_26
prod_lam_mammal_26<-calc(pgr_mammal_26, fun=prod)
plot(prod_lam_mammal_26)
```

```{r}
################function for doing it all spatially
#birds rcp 8.5
x = c(1:150)

funb<-function(y){
  
  z<-matrix(y)
  if (sum(is.na(z))<1){
    g<-mgcv:::gam(matrix(z) ~ s(x, k=8), fx=TRUE)
    g_pred<-predict(g)
    g_pred<-g_pred[55:148]
    #g_pred<-g_pred[55:100] #2050
    d<-diff(g_pred)
    gam_diff<-scale(d, center = centre_tempb, scale = scale_tempb)
    ave_mod<-data.frame(gam_diff, rep(0, length(gam_diff)), rep(0, length(gam_diff)))
    colnames(ave_mod)<-c("mean_slope_scale", "change_rate_scale", "Bodymass_scale")
    pop_pred<-predict(m1c_b, ave_mod, re.form=NA, se.fit=T)
  } else{
    
    #pop_pred$fit<-rep(NA,45) #2050
    pop_pred$fit<-rep(NA,93)
  }
  return(pop_pred$fit)
}

#gam_pred_rast_2050<-calc(obs_mod_ann, fun)
#gam_pred_rast<-calc(obs_mod_ann, fun)
```

```{r}
pred_rast_bird<-calc(obs_mod_ann, funb)

ave_lam_b<-calc(pred_rast_bird, fun=mean)
plot(10^ave_lam_b)

pgr_b<-10^pred_rast_bird
prod_lam_b<-calc(pgr_b, fun=prod)
plot(prod_lam_b)


```

```{r}
pred_rast_bird_26<-calc(obs_mod_ann_26, funb)

ave_lam_b_26<-calc(pred_rast_bird_26, fun=mean)
plot(10^ave_lam_b_26)

pgr_b_26<-10^pred_rast_bird_26
prod_lam_b_26<-calc(pgr_b_26, fun=prod)
plot(prod_lam_b_26)


```


#Future Spatial Predictions

```{r}

pop_index_2100<-function(x){10^cumsum(x)}

index_2100_bird<-calc(pred_rast_bird, pop_index_2100)
index_2100_mammal<-calc(pred_rast_mammal, pop_index_2100)

```

```{r}
library(RColorBrewer)
library(rgdal)
library(ggplot2)
library(rgeos)
world<-readOGR(dsn=getwd(), layer="ne_10m_land") #download from http://www.naturalearthdata.com/downloads/10m-physical-vectors/10m-land/
e<-c(-180, 180, -63, 83.6341)
world_c<-crop(world, e)
plot(world_c)
world_df<-fortify(world_c)
colnames(world_df)[c(1:2)]<-c("Longitude", "Latitude")

```

```{r}
pred_2100_bird<-rasterToPoints(index_2100_bird)
#pred_2100_df<-rasterToPoints(log10(pred_2100))
pred_2100_bird2<-data.frame(pred_2100_bird)
colnames(pred_2100_bird2)<-c("Longitude", "Latitude", "Population_Decline")

```

```{r}
theme_opts<-list(theme(panel.grid.minor = element_blank(),
                       panel.grid.major = element_blank(),
                       panel.background = element_rect(fill = 'white', colour = NA),
                       plot.background = element_rect(),
                       axis.line = element_blank(),
                       axis.text.x = element_blank(),
                       axis.text.y = element_blank(),
                       axis.ticks = element_blank(),
                       axis.title.x = element_blank(),
                       axis.title.y = element_blank()))

ggplot(data=pred_2100_bird2, aes(Longitude, Latitude))+
  geom_raster(aes(fill=Population_Decline))+
  scale_fill_gradientn(colors=brewer.pal(7, "RdYlGn"), guide=guide_colourbar(title=""))+ 
  #labels=c("-99%", "-50%", "0%", "+50%", "+100%"), breaks=c(0.01,0.5, 1, 1.5, 2))+   
  #labels=c("-99.9999%", "-99.99%", "-99%","0%"))+
  geom_path(data=world_df, aes(Longitude, Latitude, group=group))+
  coord_fixed(ratio = 1)+
  theme_bw()+
  labs(title="Bird Population Index 2100 - (2005 = 1)")+
  theme(text = element_text(size=20))

```


```{r}
pred_2100_mammal<-rasterToPoints(index_2100_mammal)
#pred_2100_df<-rasterToPoints(log10(pred_2100))
pred_2100_mammal2<-data.frame(pred_2100_mammal)
colnames(pred_2100_mammal2)<-c("Longitude", "Latitude", "Population_Decline")

```

```{r}
theme_opts<-list(theme(panel.grid.minor = element_blank(),
                       panel.grid.major = element_blank(),
                       panel.background = element_rect(fill = 'white', colour = NA),
                       plot.background = element_rect(),
                       axis.line = element_blank(),
                       axis.text.x = element_blank(),
                       axis.text.y = element_blank(),
                       axis.ticks = element_blank(),
                       axis.title.x = element_blank(),
                       axis.title.y = element_blank()))

ggplot(data=pred_2100_mammal2, aes(Longitude, Latitude))+
  geom_raster(aes(fill=Population_Decline))+
  scale_fill_gradientn(colors=brewer.pal(7, "RdYlGn"), guide=guide_colourbar(title=""))+ 
  #labels=c("-99%", "-50%", "0%", "+50%", "+100%"), breaks=c(0.01,0.5, 1, 1.5, 2))+   
  #labels=c("-99.9999%", "-99.99%", "-99%","0%"))+
  geom_path(data=world_df, aes(Longitude, Latitude, group=group))+
  coord_fixed(ratio = 1)+
  theme_bw()+
  labs(title="Mammal Population Index 2100 - (2005 = 1)")+
  theme(text = element_text(size=20))

```
