title: "Demoniche Workflow"
author: "Fiona Spooner"
date: "September 6, 2017"
output: html_document
---
###Chapter 3 workflow

```{r, eval=FALSE}
install.packages('climates',,'http://www.rforge.net/')
install.packages("dismo")
install.packages("maps")
install.packages("mapdata")
```

```{r, message= FALSE}
library(zoo)
library(dismo)
```

#####Creating Bioclim varibles for each year 1950-2016 for Europe - this only needs to be done once
```{r, eval=FALSE, message=FALSE}
rr<-brick(paste(wd, "/rr_0.25deg_reg_v15.0.nc", sep="")) #precipitation
tg<-brick(paste(wd, "/tg_0.25deg_reg_v15.0.nc", sep="")) #mean temp
tn<-brick(paste(wd, "/tn_0.25deg_reg_v15.0.nc", sep="")) #min temp
tx<-brick(paste(wd, "/tx_0.25deg_reg_v15.0.nc", sep="")) #max temp
#dates<-as.Date((gsub("X", "",names(rr))), format="%Y.%m.%d")


rr_mon<-zApply(rr, by=as.yearmon, fun = mean)
tg_mon<-zApply(tg, by=as.yearmon, fun = mean)
tn_mon<-zApply(tn, by=as.yearmon, fun = mean)
tx_mon<-zApply(tx, by=as.yearmon, fun = mean)

jans<-seq(1,804, by=12)
years<-as.character(1950:2016)

for (i in 1:length(years)){
  jan_sel<-jans[i]
  dec_sel<-jan_sel+11
  bio_vars_all<-biovars(rr_mon[[jan_sel:dec_sel]], tn_mon[[jan_sel:dec_sel]], tx_mon[[jan_sel:dec_sel]])  
  writeRaster(bio_vars_all, paste("C:/Users/Fiona/Documents/PhD/PhD_Method/Bioclim/", years[i],"_bioclim_variable_stack.tif", sep=""), overwrite=T)
  print(years[i])
  }
```

#####Dismo SDMs

Extracting Alpine ibex data from GBIF and adding in the locations from the LPI dataset - need to check the LPI populations are from after 1985. Formatting these points as a spatialpointsdataframe
```{r, cache=TRUE, message= FALSE, warning=FALSE}
library(dismo)
library(demoniche)
library(maps)
library(mapdata)

#Extracting Alpine ibex data from GBIF
wd<- getwd()
capra <-gbif("capra", "ibex", geo=FALSE)
capgeo <- subset(capra, !is.na(lon) & !is.na(lat) & (lon!="NA" & lat !="NA") | year!="NA") 
dups <- duplicated(capgeo[, c("lon", "lat")])
capg <-capgeo[!dups, ]
capg2 <- capg[capg$lon > 0 & capg$lon<25 & capg$lat > 43 & capg$lat < 47.9 & capg$year>=1985, ] 
capg2<-data.frame(capg2$lon,capg2$lat)
colnames(capg2)<-c("Longitude", "Latitude")

species<-"Capra_ibex"
df2<-read.csv("LPI_pops_20160523_edited.csv")
pyr<-subset(df2, Binomial ==species & Specific_location==1)    #record 11470 had wrong longitude - in Russia!
pyrs<-pyr[,c("Longitude","Latitude")]
capg2<-rbind(capg2, pyrs)
capg2<-na.omit(capg2)
capg2$presence<-rep(1)
capg2$ID<-1:nrow(capg2)

capc<-as.matrix(capg2[ , c( "Longitude","Latitude", "presence")])
capc<-matrix(capc[complete.cases(capc)], ncol=3)
xy<-as.matrix(capc[,c(1,2)])
df<-data.frame(capc[,3])
sp<-SpatialPointsDataFrame(coords=xy, data=df)
x <- circles(sp, d=50000, lonlat=TRUE)
pol <- polygons(x)
```

```{r}
e<-extent(sp)
e_big<-e+4
maps:::map('world',c('Italy', 'Switzerland', 'Austria', 'Monaco','Liechtenstein','Germany', 'Slovenia', 'France'),  col="light grey", fill=T, xlim=c(e_big[1],e_big[2]), ylim=c(e_big[3],e_big[4]))
plot(pol,add=T)
points(sp, col="red", pch=20)

```

Creating the 1985-2016 average for each bioclim variable and pulling out the layers we want to use in the model 
```{r,cache=TRUE, message=FALSE}
e<-extent(sp)+4

lf<-list.files(paste(wd, "/Bioclim/", sep=""))

first<-which(grepl("1985_bioclim", lf))
last<-which(grepl("2016_bioclim", lf))
all_years<-stack(paste(wd, "/Bioclim/",lf[first:last], sep=""))

bios<-seq(1,nlayers(all_years), by=19)

#creating a 1985-2016 average of each bioclim variable
for (i in 1:19){
  layers<-bios
  bio_layer<-mean(all_years[[layers]])
  writeRaster(bio_layer, paste(wd, "/Bioclim/Bio_",i,"_1985_2016_average.tif",sep=""), overwrite=TRUE)
  bios<-bios+1
  #print(layers)
}

#pulling out the bioclim layers i'm interested in
bio_layer_pred<-c(1,5,6,13,15,18,19)  #picking out the bioclim layers we want to use in the model
pred_nf<-stack(paste(wd, "/Bioclim/Bio_", bio_layer_pred,"_1985_2016_average.tif",sep="" ))

```

Creating training and testing presence and background data
```{r, cache=TRUE, message=FALSE}
k<-4
group_pres<-kfold(sp, k)
pres_train<-sp[group_pres!=1,]
pres_test<-sp[group_pres==1,]

set.seed(10)
backg <- randomPoints(pred_nf, n=1000, ext=e, extf = 1.25)
colnames(backg) = c('lon', 'lat')
group_back <- kfold(backg, k)
backg_train <- backg[group_back != 1, ]
backg_test <- backg[group_back == 1, ]

pres_train<-data.frame(pres_train)[,c(2,3)]
colnames(pres_train)<-c("lon", "lat")
train <- rbind(pres_train, backg_train)

pres_test<-data.frame(pres_test)[,c(2,3)]
colnames(pres_test)<-c("lon", "lat")
test<-rbind(pres_test, backg_test)

pb_train <- c(rep(1, nrow(pres_train)), rep(0, nrow(backg_train)))
envtrain <- extract(pred_nf, train)
envtrain <- data.frame( cbind(pa=pb_train, envtrain) )

pb_test<-c(rep(1, nrow(pres_test)), rep(0, nrow(backg_test)))
envtest <- extract(pred_nf, test)
envtest <- data.frame( cbind(pa=pb_test, envtest) )

env_all<-rbind(envtrain, envtest)
```

#####Bioclim

Running the bioclim envelope model and evaluating it to get an AUC value

```{r, cache=TRUE}
#Bioclim
bc <- bioclim(pred_nf, sp)
ev <- dismo:::evaluate(pres_test, backg_test, bc, pred_nf)

group_back <- kfold(backg, k)

#group_back <- kfold(pb_train, k)
evl<- list()

for (i in 1:k){
  pres_train<-sp[group_pres!=i,]
  pres_test<-sp[group_pres==i,]
  #backg_train<-sp[group_back!=i,]
  backg_test<-backg[group_back==i,]
  bc <- bioclim(pred_nf,pres_train)
  evl[[i]] <- dismo:::evaluate(pres_test, backg_test, bc, pred_nf)
  #print(i)
}

auc <- sapply( evl, function(x){slot(x, "auc")} )
print(auc)
bioclim_auc<-mean(auc)
print(bioclim_auc)
```

#####GAM

Running a gam and evaluating it to get an AUC value

```{r, cache= TRUE, message=FALSE}
library(mgcv)
gm1<-gam(pa~ s(Bio_1_1985_2016_average)+s(Bio_5_1985_2016_average)+ s(Bio_6_1985_2016_average)+ 
           s(Bio_13_1985_2016_average)+ s(Bio_15_1985_2016_average)+ s(Bio_18_1985_2016_average)+ 
           s(Bio_19_1985_2016_average), family = binomial(link = "logit"),data=env_all)

gp<-predict(pred_nf, gm1)
r<-calc(gp, fun=function(x){ exp(x)/(1+exp(x))})

env_pres<-env_all[env_all$pa==1,]
env_back<-env_all[env_all$pa==0,]
group_pres<-  kfold(env_pres, k)
group_back<- kfold(env_back, k)

evl<- list()

for (i in 1:k){
  pres_train<-env_pres[group_pres!=i ,]
  pres_test<-env_pres[(group_pres==i) ,]
  back_test<-env_back[(group_back==i),]
  back_train<-env_back[(group_back!=i),]
  envtrain<-rbind(pres_train, back_train)
  gm1<-gam(pa~ s(Bio_1_1985_2016_average)+ s(Bio_5_1985_2016_average)+ 
             s(Bio_6_1985_2016_average)+ s(Bio_13_1985_2016_average)+ 
             s(Bio_15_1985_2016_average)+ s(Bio_18_1985_2016_average)+ 
             s(Bio_19_1985_2016_average),family= binomial(link = "logit") , data=envtrain)
  evl[[i]] <- dismo:::evaluate(pres_test, back_test, gm1)
  #print(i)
}

auc <- sapply( evl, function(x){slot(x, "auc")} )
print(auc)

gam_auc<-mean(auc)
print(gam_auc)

```

#####Random Forest

Running the random forestenvelope model and evaluating it to get an AUC value

```{r, cache= TRUE, message= FALSE, warning=FALSE}
library(randomForest)

model<-pa~  Bio_1_1985_2016_average+Bio_5_1985_2016_average+ 
  Bio_6_1985_2016_average+ Bio_13_1985_2016_average+ Bio_15_1985_2016_average+ 
  Bio_18_1985_2016_average+ Bio_19_1985_2016_average

evl<- list()

for (i in 1:k){
  pres_train<-env_pres[group_pres!=i,]
  pres_test<-env_pres[(group_pres==i) ,]
  back_test<-env_back[(group_back==i),]
  back_train<-env_back[group_back !=i,]
  envtrain<-rbind(pres_train, back_train)
  rf1 <- randomForest(model, data=envtrain)
  evl[[i]] <- dismo:::evaluate(pres_test, back_test, rf1)
}

auc <- sapply( evl, function(x){slot(x, "auc")} )
print(auc)

rf_auc<-mean(auc)
print(rf_auc)
```


#####Using the models to predict

Creating a weighted ensemble averaged suitability model for 1985-2016 which can be used to predict suitability backwards

```{r, cache=TRUE, warning=FALSE, message=FALSE}
#total models
bc <- bioclim(pred_nf, sp)
gm1<-gam(pa~ s(Bio_1_1985_2016_average)+s(Bio_5_1985_2016_average)+ s(Bio_6_1985_2016_average)+ 
           s(Bio_13_1985_2016_average)+ s(Bio_15_1985_2016_average)+ s(Bio_18_1985_2016_average)+ 
           s(Bio_19_1985_2016_average), family = binomial(link = "logit"),data=env_all)
rf1 <- randomForest(model, data=env_all)

pb <- predict(pred_nf, bc, ext=e, progress='') #bioclim predict
pg <- predict(pred_nf, gm1, ext=e) #gam predict
pgl<-calc(pg, fun=function(x){ exp(x)/(1+exp(x))}) #backtransforming from logit space
pr <- predict(pred_nf, rf1, ext=e) #random forest predict

models <- stack(pb, pgl, pr)
names(models) <- c("bioclim", "gam", "random forest")
plot(models)

auc<-c(bioclim_auc, gam_auc, rf_auc)
w <- (auc-0.5)^2
wm <- weighted.mean( models[[c("bioclim", "gam", "random.forest")]], w)
plot(wm)

```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#thresholds - perhaps redundant for now?
library(rgdal)

bc <- bioclim(pred_nf, sp)
gm1<-mgcv:::gam(pa~ s(Bio_1_1985_2016_average)+ s(Bio_5_1985_2016_average)+ s(Bio_6_1985_2016_average)+ s(Bio_13_1985_2016_average)+ s(Bio_15_1985_2016_average)+s(Bio_18_1985_2016_average)+ s(Bio_19_1985_2016_average), data=envtrain)
rf1 <- randomForest:::randomForest(model, data=envtrain)

k<-4
group_pres<-kfold(sp, k)
pres_train<-sp[group_pres!=1,]
pres_test<-sp[group_pres==1,]

set.seed(10)
backg <- randomPoints(pred_nf, n=1000, ext=e, extf = 1.25)
colnames(backg) = c('lon', 'lat')
group_back <- kfold(backg, k)
backg_train <- backg[group_back != 1, ]
backg_test <- backg[group_back == 1, ]

ev <- dismo:::evaluate(pres_test, backg_test, bc, pred_nf)
tr_bc <- threshold(ev, 'spec_sens')
tr_bc_k <- threshold(ev, 'kappa')
tr_bc_p <- threshold(ev, 'prevalence')
tr_bc_e<-threshold(ev, 'equal_sens_spec')
tr_bc_tss<-ev@t[which.max(ev@TPR + ev@TNR -1)]


gam_ev<-dismo:::evaluate(pres_test, backg_test, gm1, pred_nf)
tr_gam<-threshold(gam_ev, 'spec_sens')
tr_gam_k<-threshold(gam_ev, 'kappa')
tr_gam_p<-threshold(gam_ev, 'prevalence')
tr_gam_e<-threshold(gam_ev, 'equal_sens_spec')
tr_gam_tss<-gam_ev@t[which.max(gam_ev@TPR + gam_ev@TNR -1)]

rf_ev<-dismo:::evaluate(pres_test, backg_test, rf1, pred_nf)
tr_rf<-threshold(rf_ev, 'spec_sens')
tr_rf_k<-threshold(rf_ev, 'kappa')
tr_rf_p<-threshold(rf_ev, 'prevalence')
tr_rf_e<-threshold(rf_ev, 'equal_sens_spec')
tr_rf_tss<-rf_ev@t[which.max(rf_ev@TPR + rf_ev@TNR -1)]

thresh<-mean(c(tr_bc, tr_gam, tr_rf))
thresh_weighted<-weighted.mean(c(tr_bc, tr_gam, tr_rf), w)
thresh_weighted_k<-weighted.mean(c(tr_bc_k, tr_gam_k, tr_rf_k), w)
thresh_weighted_p<-weighted.mean(c(tr_bc_p, tr_gam_p, tr_rf_p), w)
thresh_weighted_e<-weighted.mean(c(tr_bc_e, tr_gam_e, tr_rf_e), w)
thresh_weighted_tss<-weighted.mean(c(tr_bc_tss, tr_gam_tss, tr_rf_tss))

tholds<-c(thresh, thresh_weighted, thresh_weighted_k, thresh_weighted_p, thresh_weighted_e, thresh_weighted_tss)
names<-c("non_weighted", "spec_sens", "kappa", "prevalence", "equal_sens_spec", "TSS")

cbind(names, tholds)
```

Using the ensemble model to predict suitability for each year - 1950-2016

```{r, cache=TRUE, eval=FALSE}
years<-1950:2016
library(raster)
library(mgcv)
library(randomForest)
for (i in 1:length(years)){
  
  pred_nf<-stack(paste(wd, "/Bioclim/", years[i], "_bioclim_variable_stack.tif", sep="" ))  
  pred_nf<-pred_nf[[bio_layer_pred]]
  names(pred_nf)<-c("Bio_1_1985_2016_average","Bio_5_1985_2016_average", "Bio_6_1985_2016_average", "Bio_13_1985_2016_average" ,"Bio_15_1985_2016_average","Bio_18_1985_2016_average", "Bio_19_1985_2016_average")
  
  pb <- predict(pred_nf, bc, ext=e, progress='')
  pg <- predict(pred_nf, gm1, ext=e) #gam predict
  pgl<-calc(pg, fun=function(x){ exp(x)/(1+exp(x))})
  pr <- predict(pred_nf, rf1, ext=e)
  
  models <- stack(pb, pgl, pr)
  names(models) <- c("bioclim", "gam", "random forest")
  wm <- weighted.mean( models[[c("bioclim", "gam", "random.forest")]], w)
  
  pa<-wm>thresh_weighted
  pa_k<-wm>thresh_weighted_k
  pa_p<-wm>thresh_weighted_p
  pa_e<-wm>thresh_weighted_e
  pa_tss<- wm>thresh_weighted_tss
  
 #writeRaster(wm , paste(wd, "/Alp_SDMs/Ensembles/weighted_ensemble_sdm_", years[i], ".tif", sep=""), overwrite=TRUE)
  #writeRaster(pa , paste(wd, "/Alp_SDMs/Ensembles/pres_abs_weighted_ensemble_sdm_", years[i], ".tif", sep=""), overwrite=TRUE)
  #writeRaster(pa_k , paste(wd, "/Alp_SDMs/Ensembles/pres_abs_kappa_weighted_ensemble_sdm_", years[i], ".tif", sep=""), overwrite=TRUE)
  #writeRaster(pa_p , paste(wd, "/Alp_SDMs/Ensembles/pres_abs_prevalence_weighted_ensemble_sdm_", years[i], ".tif", sep=""), overwrite=TRUE)
  #writeRaster(pa_e , paste(wd, "/Alp_SDMs/Ensembles/pres_abs_equal_sens_spec_weighted_ensemble_sdm_", years[i], ".tif", sep=""), overwrite=TRUE)
  writeRaster(pa_tss, paste(wd, "/Alp_SDMs/Ensembles/pres_abs_true_skill_stat_weighted_ensemble_sdm_", years[i], ".tif", sep=""), overwrite=TRUE )
  
  print(years[i])
  # print(cellStats(pa, "max"))
  # print(cellStats(pa_k, "max"))
  # print(cellStats(pa_p, "max"))
  # print(cellStats(pa_e, "max"))
  #plot(wm, main=years[i])
}

```
#Copying the 1950 files for the 900 year spin up
```{r, eval=FALSE}
spin_years<-1049:1949
from_sdm<- paste(wd, "/Alp_SDMs/Ensembles/weighted_ensemble_sdm_1950.tif", sep="")
to_sdm<-paste(wd, "/Alp_SDMs/Ensembles/weighted_ensemble_sdm_",spin_years, ".tif", sep="")

from_pa<- paste(wd, "/Alp_SDMs/Ensembles/pres_abs_weighted_ensemble_sdm_1950.tif", sep="")
to_pa<-paste(wd, "/Alp_SDMs/Ensembles/pres_abs_weighted_ensemble_sdm_",spin_years, ".tif", sep="")

from_kap<- paste(wd, "/Alp_SDMs/Ensembles/pres_abs_kappa_weighted_ensemble_sdm_1950.tif", sep="")
to_kap<-paste(wd, "/Alp_SDMs/Ensembles/pres_abs_kappa_weighted_ensemble_sdm_",spin_years, ".tif", sep="")

from_prev<-paste(wd, "/Alp_SDMs/Ensembles/pres_abs_prevalence_weighted_ensemble_sdm_1950.tif", sep="")
to_prev<-paste(wd, "/Alp_SDMs/Ensembles/pres_abs_prevalence_weighted_ensemble_sdm_",spin_years, ".tif", sep="")

from_eq<-paste(wd, "/Alp_SDMs/Ensembles/pres_abs_equal_sens_spec_weighted_ensemble_sdm_1950.tif", sep="")
to_eq<-paste(wd, "/Alp_SDMs/Ensembles/pres_abs_equal_sens_spec_weighted_ensemble_sdm_",spin_years, ".tif", sep="")

file.copy(from_sdm,to_sdm)
file.copy(from_pa,to_pa)
file.copy(from_kap,to_kap)
file.copy(from_prev,to_prev)
file.copy(from_eq,to_eq)


```

### Demoniche

Formatting the data for use in demoniche
```{r, warning=FALSE, message=FALSE}
library(demoniche)
library(rgdal)
library(raster)
species<-"Capra_ibex"

df2<-read.csv("LPI_pops_20160523_edited.csv")
pyr<-subset(df2, Binomial ==species & Specific_location==1)    #record 11470 had wrong longitude - in Russia!
# tl<-pyr[,c(65:130)]
# tl[tl=="NULL"]<-NA
# tlm<-as.matrix(tl)
# tlmn<-as.numeric(tlm)
# tl<-matrix(tlmn, nrow=10, ncol=66)
# ann_sum<-colSums(tl, na.rm=T)
# plot(1950:2015, ann_sum)

plot(pyr$Longitude, pyr$Latitude)

#formatting the data for use in demoniche
pyrs<-pyr[,c("ID","Longitude","Latitude")]

id<-pyrs$ID*100
lam<-rep(1,length(id))    #not sure what the value here pertains to - think it sets starting population so should use values from LPI?
pyrxy<-SpatialPoints(pyr[,c("Longitude","Latitude")])

wd<-getwd()
sdm<-raster(paste(wd,"/Alp_SDMs/Ensembles/pres_abs_weighted_ensemble_sdm_1950.tif", sep=""))
e2<-extent(sdm)

r<-raster(e2, resolution=res(sdm))

rz<-rasterize(pyrxy,r,lam )
rid<-rasterize(pyrxy,r,id)
# plot(rz)
# plot(rid)

rz_spdf<-xyFromCell(rz, 1:ncell(rid))

rzm<-as.vector(rz)
ridm<-as.vector(rid)

df<-data.frame(ridm,rz_spdf,rzm)
colnames(df)<-c( "PatchID","X","Y","area")

Populations<-data.frame(na.omit(df)) 

```


```{r, cache=TRUE, echo=FALSE}

#formatting the sdm data


years<-1050:2016

sdm_patch_df<-data.frame(ID=1:ncell(rid))
sdm_df<-data.frame(ID=1:ncell(rid))


#formatting data for demoniche
for (i in 1:length(years)){
  
  #a selection of different threshold techniques for presence absence, as well as a suitability surface
  sdm<-raster(paste(wd,"/Alp_SDMs/Ensembles/weighted_ensemble_sdm_", years[i],".tif", sep=""))   #14.6
  #sdm<-raster(paste(wd,"/Alp_SDMs/Ensembles/pres_abs_weighted_ensemble_sdm_", years[i],".tif", sep="")) #15.4%  
  #patch<-raster(paste(wd,"/Alp_SDMs/Ensembles/pres_abs_kappa_weighted_ensemble_sdm_", years[i],".tif", sep="")) #21.1%  
  #sdm<-raster(paste(wd,"/Alp_SDMs/Ensembles/pres_abs_prevalence_weighted_ensemble_sdm_", years[i],".tif", sep="")) #14,3%  
  #sdm<-raster(paste(wd,"/Alp_SDMs/Ensembles/pres_abs_equal_sens_spec_weighted_ensemble_sdm_", years[i],".tif", sep=""))  #13.1%
  patch<-raster(paste(wd, "/Alp_SDMs/Ensembles/pres_abs_true_skill_stat_weighted_ensemble_sdm_", years[i],".tif", sep=""))
  if (i ==1){
    vec<-as.data.frame(sdm, xy = TRUE)
    vec_pat<-as.data.frame(patch, xy=TRUE)
  } else{
    vec<-as.data.frame(sdm)
    vec_pat<-as.data.frame(patch)
  }
  
  sdm_df<-cbind(sdm_df, vec)
  sdm_patch_df<-cbind(sdm_patch_df,vec_pat)
  #print(i)
}

sdm_df$ID[which(!is.na(df$PatchID))]<-df$PatchID[!is.na(df$PatchID)]
sdm_patch_df$ID[which(!is.na(df$PatchID))]<-df$PatchID[!is.na(df$PatchID)]

niche_map_mine<-sdm_df
colnames(niche_map_mine)[1:3]<-c("gridID", "X", "Y")


patch_map_mine<-sdm_patch_df
colnames(patch_map_mine)[1:3]<-c("gridID", "X", "Y")

col_years<-paste("Year_", 1950:2016, sep="")
col_years<-paste("Year_", 1050:2016, sep="")

colnames(niche_map_mine)[4:length(colnames(niche_map_mine))]<-col_years
colnames(patch_map_mine)[4:length(colnames(patch_map_mine))]<-col_years

#niche_formulas <- as.formula(paste(paste(colnames(niche_map_mine)[-c(1:3)], collapse="+"),"X+Y",sep="~"))

#print(levelplot(niche_formulas, niche_map_mine, col.regions=rev(heat.colors(100)), main = "Niche Values"))

no_yrs_mine<-1 #number of years each time period represents 


#plot(years,colMeans(na.omit(niche_map_mine)[4:length(colnames(niche_map_mine))]), type="l", ylab="Mean suitability index")

```

#####Comadre

Accessing the population matrices from comadre

```{r, warning=FALSE, message=FALSE}
#Population matrix set up


load(paste(wd, "COMADRE_v.2.0.1.RData", sep="/"))

#load(paste(wd, "COMADRE_v.1.0.0.RData", sep="/"))

species<-"Capra_ibex"

tempMetadata<-subset(comadre$metadata, SpeciesAccepted==species)

keep<-as.numeric(rownames(tempMetadata))

tempMat<-comadre$mat[keep]   #MatA is matrix pop model, can be split into U, F and/or C

MatList<-list(tempMat[[1]][[1]])  #varies depending on number of matrices - need to find a way to code this better - now have five matrices available so need to sort this
AllMat<-unlist(MatList)
matrices<-matrix(AllMat, ncol=length(MatList))
colnames(matrices)<- c("Reference_matrix")

#tempMat<-(tempMat[[2]][[1]] + tempMat[[3]][[1]] + tempMat[[4]][[1]]+ tempMat[[5]][[1]])/4
#MatList<-list(tempMat[[2]][[1]], tempMat[[3]][[1]] ,tempMat[[4]][[1]], tempMat[[5]][[1]])
# AllMat<-unlist(MatList)
# matrices<-matrix(AllMat, ncol=length(MatList))
# #matrices<-matrix(tempMat, ncol=1)
# colnames(matrices)<- c("Reference_matrix", "Matrix_1", "Matrix_2", "Matrix_3")

```

Parameterising the demoniche model

```{r, cache=TRUE}
prob_scenario<-c(0.5,0.5)    #need to check this

noise<-0.90     #not yet active in demoniche

stages<-comadre$matrixClass[keep][[1]]$MatrixClassAuthor
#stages<-comadre$matrixClass[keep][[2]]$MatrixClassAuthor
#stagesf<-stages[1:3]

list_names_matrices<-colnames(matrices)

sumweight<-rep(1, length(stages)) #weight of stages  - should be equal for all mine just 
#in plants seed not included in calculating population sizes - or if you wanted to just 
#calculate the female population it would be c(1,1,1,0,0,0)
#sumweightf<-c(1,1,1)

transition_affected_niche<-c(2,11,20)   #which parts of the matrix are affected by the c(1,2) is juveniles
#niche values

transition_affected_env <- "all"

transition_affected_demogr <- "all"

env_stochas_type<-"normal"   #can also be lognormal

matrices_var <- matrix(0.01, ncol = 1, nrow = nrow(matrices), dimnames = list(NULL, "sd")) 
#standard deviation of matrices

proportion_initial<- rep(1/length(stages), length(stages)) #proportion of population in 
#each stage - no idea what this should be and will likely have a big impact on results! 
#- just doing eqaul splits for now
#proportion_initialf<- c(1/3,1/3,1/3)

density_individuals <- 4000  #4292.32 to 16096.2 based on density being between 8 and 30 per 100 ha and the area of each cell being 53654 ha - 


K_weight<-c(rep(1, length(stages)))  #the weight with which carrying capacity affects each stage was FALSE


#niche_map_mine[is.na(niche_map_mine)]<-0
patch_map_mine[is.na(patch_map_mine)]<-0
```

#####Demoniche setup

Running the demoniche model

```{r, cache=TRUE}
#plotting functions linking k and hsi

k<-1 #set to one when using filzbach estimates - otherwise set to a standard value across all pops
lin<-function(x){
  x*k
  }

sig<-function(x){
  k*(1/(1+exp(-10*x+5)))
  }   #could also try with 18x + 9 - closer to SM in Damaris paper

lt<-function(x) { 
  val = (4/3) * x - (1/3)
  val[x < 0.25] = 0 
  val<-val*k
  return(val)
}


#carrying capacity function 

lf<-list.files(paste(wd,"/Alp_SDMs/Ensembles/", sep=""))   

files<-lf[grepl("^weighted_ensemble_sdm_.*.tif$", lf)]

sdms<-stack(paste(wd,"/Alp_SDMs/Ensembles/", files, sep=""))

ks_df<-read.csv("carrying_capacity_log_likelihood_capra_ibex.csv")
ks_df$PatchID<-ks_df$ID*100

#for 1950-2016
#sdms<-sdms[[901:967]]




ks_xy<-merge(Populations, ks_df[,c(2:4)], by="PatchID")


hsi<-extract(sdms, ks_xy[,c(2,3)])


link<-lin(hsi)
links<-link*ks_xy$ks

sigk<-sig(hsi)
sigks<-sigk*ks_xy$ks

ltk<-lt(hsi)
ltks<-ltk*ks_xy$ks

plot(links[1,c(900:967)], type="l")

```


```{r}

library(demoniche)
source("demoniche_model_me.R")


foldernames<-c("patch_link_low_disp" ,"patch_link_mid_disp" ,"patch_link_high_disp","patch_sigk_low_disp" ,"patch_sigk_mid_disp" ,"patch_sigk_high_disp","patch_ltk_low_disp" ,"patch_ltk_mid_disp" ,"patch_ltk_high_disp" )


k_relationships<-rep(list(links, sigks, ltks), each=3)  #need to change names of folders to match this

dispersal_levels<-rep(c(0.1,0.5,0.9), each=3)

for (i in 2:length(foldernames)){

  print(paste("processing of ", foldernames[i]," beginning, ", i, " of 9", sep=""))
  foldername = foldernames[i]
  dispersal = dispersal_levels[i]
  K = matrix(unlist(k_relationships[i]),nrow=9)
  
  demoniche_setup(modelname = foldername,Populations = Populations, Nichemap =patch_map_mine,
                matrices = matrices, matrices_var = matrices_var, prob_scenario = prob_scenario,
                stages = stages, proportion_initial = proportion_initial,
                density_individuals = density_individuals,
                fraction_LDD = dispersal, 
                fraction_SDD = dispersal,
                dispersal_constants = c(0.7,0.7,0.1,1),
                transition_affected_demogr = transition_affected_demogr,
                transition_affected_env=transition_affected_env,
                env_stochas_type = env_stochas_type,
                no_yrs = no_yrs_mine, K=K, Kweight = K_weight, 
                sumweight =sumweight)

patch_run <- demoniche_model_me(modelname = foldernames[i], Niche = TRUE, 
                                     Dispersal = TRUE, repetitions = 1,
                                     foldername = foldernames[i])
years<-1050:2016
plot(years,patch_run[,"Meanpop","Reference_matrix"], type="l", xlim=c(1949,2016), main=foldernames[i])

}

```


Converting pop output
```{r}



convert_pop_out<-function(foldername){
  
  pop_out<-read.csv(paste(wd, "/", foldername, "/pop_output.csv", sep=""), header = TRUE)
  pop_out<-pop_out[,-1]
  coordinates(pop_out) <- ~ X + Y
  gridded(pop_out) <- TRUE
  rasterDF <- stack(pop_out)
  trends<-extract(rasterDF,xy)
  trends_df<-data.frame(trends)
  write.csv(trends_df, paste(wd, "/", foldername, "/pop_trend_output.csv", sep=""))
}

lapply(foldernames, convert_pop_out)


```

```{r, eval=FALSE}
lpi<-read.csv("LPI_pops_20160523_edited.csv")

alpine<-lpi[lpi$ID == 10696 | lpi$ID == 10714 | lpi$ID == 10694 | lpi$ID == 10717 | lpi$ID == 10713 |lpi$ID == 10718 |lpi$ID == 10695 |lpi$ID == 539 |lpi$ID == 10710 ,]

pop<-3

plot(years, trends[pop,], type="l", ylim=c(0,5000), xlim=c(1950,2016))
lpi_pop<-alpine[,c(65:130)]
lines(years[901:966], lpi_pop[pop,], type="l", col="red")

pop_trends_pred<-data.frame(cbind(Populations$PatchID, trends))
library(reshape)
melt_pop<-melt(pop_trends_pred)

melt_pop<-melt_pop[-c(1:9),]

lpi_pop<-data.frame(cbind(Populations$PatchID, lpi_pop))
lpi_pop$Populations.PatchID<-as.character(lpi_pop$Populations.PatchID)
melt_lpi<-melt(lpi_pop, id.vars = "Populations.PatchID")


melt_pop$ID<-as.factor(lpi_pop$Populations.PatchID)
melt_pop$Year<-rep(1050:2016 , each=9)

melt_pop_1950<-subset(melt_pop, melt_pop$Year >=1950)

library(ggplot2)
#plot of predicted trends
ggplot(data = melt_pop_1950, aes(y= value, x= Year, group=ID, col=ID))+
  geom_line()

col_years<-paste("Year_", 1950:2015, sep="")
colnames(lpi_pop)<-col_years


melt_lpi$Year<-rep(1950:2015 , each=9)
melt_lpi$value<-as.numeric(as.character(melt_lpi$value))

#plot of actual trends
ggplot(data = melt_lpi, aes(y= value, x= Year, group=Populations.PatchID, col=Populations.PatchID))+
  geom_line()

```

```{r}
library(lme4)
dt<-read.csv("Mammals_scaled_ready_for_models.csv")

m1b<-lmer(lambda_mean ~ change_rate_scale+(1|Binomial)+(1|loc_id),data=dt, REML=F)
m1c<-lmer(lambda_mean ~ mean_slope_scale+(1|Binomial)+(1|loc_id),data=dt)


# 
# b1s<-coef(m1c)
# 
# icol <- which(colnames(dt)=="mean_slope_scale")
# p.order <- c(icol,(1:ncol(dt))[-icol])
# m <- c(0,mean(dt$Estimate))
# s <- c(1,sd(dt$Estimate))
# all.equal(m1c,rescale.coefs(b1s,m,s)) 


capra_df<-dt[dt$ID == 10696 | dt$ID == 10714 |dt$ID == 10694 | dt$ID == 10717 | dt$ID == 10713 |dt$ID == 10718 |dt$ID == 10695 |dt$ID == 539 |dt$ID == 10710 ,]


capras<-which(dt$ID == 10696 | dt$ID == 10714 | dt$ID == 10694 | dt$ID == 10717 | dt$ID == 10713 |dt$ID == 10718 |dt$ID == 10695 |dt$ID == 539 |dt$ID == 10710 )


pred<-10^(predict(m0)[capras])
pred_clim<-10^predict(m1c)[capras]   #climate only model
pred_luc<-10^predict(m1b)[capras]   #land use only model
real<-10^capra_df$lambda_mean
  
demoniche<-10^rowMeans(diff(as.matrix(log10(trends_df[,c(900:967)]))))


lambda_preds<-cbind(pred, pred_clim,pred_luc, real, demoniche)


pairs(lambda_preds)


```


```{r, eval=FALSE, echo=FALSE}

patch_run_dz <- demoniche_model_VE(modelname = "Capra_ibex_patch", Niche = TRUE, 
                                     Dispersal = T, repetitions = 1,
                                     foldername = "damaris_function")

patch_run_dz[,"Meanpop","Reference_matrix"]

years<-1050:2016
#plot(RPyran_disp_niche[,"Meanpop","Reference_matrix"])
bleh<-(patch_run[,"Meanpop","Reference_matrix"])

plot(years, bleh, type="l", xlim=c(1950,2016))

```


HSM - linear k - all stages
```{r, eval=FALSE, echo=FALSE}

demoniche_setup(modelname = "Capra_ibex_scale_link",Populations = Populations, Nichemap = niche_map_mine,
              matrices = matrices,matrices_var = matrices_var, prob_scenario = prob_scenario,
                stages = stages, proportion_initial = proportion_initial,
                density_individuals = 400,  #number of individuals at the start of each population - can be a vector of length(Populations)
                fraction_LDD = 0.1, fraction_SDD = 0.9,
                dispersal_constants = c(0.7,0.7,0.1,2),
                transition_affected_niche = "all",
                transition_affected_demogr = transition_affected_demogr,
                transition_affected_env=transition_affected_env,
                env_stochas_type = env_stochas_type,
                no_yrs = no_yrs_mine, K=link, Kweight = K_weight, 
                sumweight =sumweight)

RPyran_disp_niche_scale_link <- demoniche_model(modelname = "Capra_ibex_scale_link", Niche = TRUE, 
                                     Dispersal = TRUE, repetitions = 1,
                                     foldername = "non_damaris_function")

RPyran_disp_niche_scale <- demoniche_model(modelname = "Capra_ibex_scale", Niche = TRUE, 
                                     Dispersal = TRUE, repetitions = 1,
                                     foldername = "RPyran_disp_niche")


years<-1950:2016
years<-1050:2016
#plot(RPyran_disp_niche[,"Meanpop","Reference_matrix"])
bleh<-(RPyran_disp_niche_scale[,"Meanpop","Reference_matrix"])
link_years<-(RPyran_disp_niche_scale_lk[,"Meanpop","Reference_matrix"])
lk_years<-(RPyran_disp_niche_scale_link[,"Meanpop","Reference_matrix"])
#plot(years, bleh, type="l", ylab="Number of Alpine Ibex")


plot(years[900:967], bleh[900:967], type="l", ylab="Number of Alpine Ibex", ylim=c(0,1100000), xlim=c(1950,2016), xlab="")
lines(years[900:967], lk_years[900:967], type="l", ylab="Number of Alpine Ibex", ylim=c(0,1100000), xlim=c(1950,2016), xlab="", col="red")
lines(years[900:967], link_years[900:967], type="l", ylab="Number of Alpine Ibex", ylim=c(0,1100000), xlim=c(1950,2016), xlab="", col="green")


mean(diff(log10(bleh[900:967])), type="l")
mean(diff(log10(lk_years[900:967])), type="l")
mean(diff(log10(link_years[900:967])), type="l")

```

```{r, eval=FALSE, echo=FALSE}

spin_up<-1050:1949
year0<-paste(wd, "/Alp_SDMs/Ensembles/pres_abs_true_skill_stat_weighted_ensemble_sdm_1950.tif", sep="")

spin_up_files<-paste(wd, "/Alp_SDMs/Ensembles/pres_abs_true_skill_stat_weighted_ensemble_sdm_", spin_up,".tif", sep="")


file.copy(year0, spin_up_files)



```

